Potential of I/O-Aware Workflows in Climate and Weather

### 2020-05-22 paper.tex
New additions about Cylc

The key contributions of this paper are:
1) A sketch of a vision for an integrated data-driven approach, with a discussion of the associated challenges and implications, and 2) An architecture and roadmap consistent with this vision that would allow seamless integration into current climate and weather workflows as it utilises versions of existing tools (Cylc, XIOS, DDN IME, and ESDM).

@@

Burst buffer solutions provide a tiered storage system that aims to exploit a storage hierarchy.
They can be integrated into hardware capabilities such as DDN IME~\cite{BODIAIFSFI19} or simple software solutions.
A policy system, e.g., deployed on a burst buffer~\cite{RomanusRP15}, aims to simplify the data movement for the user, but typically migrates objects in the coarse granularity of files.
File systems and data management software such as IBM Spectrum Scale~\cite{SchmuckH02}, HPSS~\cite{528214}, BeeGFS~\cite{beegfs}, and Lustre~\cite{abs-1903-01955} (e.g., using the progressive file layouts feature) provide hierarchical storage management allowing to store data on different storage technology according to administrator-provided policies.
However, the semantic information that can be used by this type of system to make decisions is limited, e.g., data location, file extension, age of the file, etc.

@@

\begin{description}[itemsep=0pt]
  \item[Name] A basic name for the field/data generated. It is extended by a pattern defined in a variable (Lines: 8, 16, 23).

@@

From the list of opportunities, we realise data placement and migration in a heterogeneous (multi-storage) environment.
These goals will be achieved via the proposed I/O-aware scheduler, called here EIOS (ESDM I/O Scheduler).
EIOS will make the schedule considering Cylc workflow, and ESDM provided system characteristics.
Components of it are involved in different steps of the workflow and the I/O path.

While Cylc schedules the workflow, EIOS can provide hints about colocating tasks which provide the opportunity for keeping data in local storage.
Decisions about data locality will not be made for a whole (and potentially big) workflow.
Instead, the system will make decisions by looking ahead to several steps of the workflow, allowing reacting to the observed dynamics of the execution.
We utilise DDN's IME API to pin data in IME and to trigger migrations between IME and a storage backend explicitly.

Ultimately, when a user-script runs, the information about the intended I/O schedule is communicated from EIOS through a modified filename, which is then used by the ESDM-aware application to determine the data placement.

>>>>

The key contributions of this paper are:
1) A sketch of a vision for an integrated data-driven approach, with a discussion of the associated challenges and implications, and 2) An architecture and roadmap consistent with this vision that would allow seamless integration into current climate and weather workflows as it utilises versions of existing tools (Cylc, XIOS, DDN's IME, and ESDM).

@@

Burst buffer solutions provide a tiered storage system that aims to exploit a storage hierarchy.
They can be integrated into hardware capabilities such as DDN's Infinite Memory Engine (IME)~\cite{BODIAIFSFI19} or simple software solutions.
A policy system, e.g., deployed on a burst buffer~\cite{RomanusRP15}, aims to simplify the data movement for the user, but typically migrates objects in the coarse granularity of files.
File systems and data management software such as IBM Spectrum Scale~\cite{SchmuckH02}, HPSS~\cite{528214}, BeeGFS~\cite{beegfs}, and Lustre~\cite{abs-1903-01955} (e.g., using the progressive file layouts feature) provide hierarchical storage management allowing to store data on different storage technology according to administrator-provided policies.
However, the semantic information that can be used by this type of system to make decisions is limited, e.g., data location, file extension, age of the file, etc.

@@

\begin{description}[itemsep=0pt]

  \item[Name] A basic name for the field/data generated. It is extended by a pattern defined in a variable (Lines: 8, 16, 23).

@@

From the list of opportunities, we realise data placement and migration in a heterogeneous (multi-storage) environment.
These goals will be achieved via the proposed I/O-aware scheduler, called here EIOS (ESDM I/O Scheduler).
EIOS will make the schedule considering Cylc workflow, and ESDM provided system characteristics.
Components of it are involved in different steps of the workflow and the I/O path.
While Cylc schedules the workflow, EIOS can provide hints about colocating tasks which provide the opportunity for keeping data in local storage.
There are two main requirements from Cylc, both covered by Cylc normal functionalities:

\item[The ability to dynamically set the job (Slurm) directives for a task]

\ \\ \vspace{-0.5cm}

This will be achieved by calling an external command (run on the Cylc scheduler host) which defines directives to be used by the job. This external command will determine some attributes of previous tasks through simple SQL queries to the Cylc database. We plan on using Cylc broadcast to alter the directives used by a task by running an extra local task prior to any task where we need to be able to alter the directives.

\item[The ability to dynamically set storage locations]

\ \\ \vspace{-0.5cm}

This will be achieved by defining environment variables in the job script which are set to the output of another external command. This external command (run on the job host) will have access to all the normal Cylc environment variables providing details about the current task.

We plan on utilising DDN's IME API to pin data in IME and to trigger migrations between IME and a storage backend explicitly.
Decisions about data locality will not be made for a whole (and potentially big) workflow.
Instead, the system will make decisions by looking ahead to several steps of the workflow, allowing reacting to the observed dynamics of the execution.
Ultimately, when a user-script runs, the information about the intended I/O schedule is communicated from EIOS through a modified filename, which is then used by the ESDM-aware application to determine the data placement.
